Synthesizing Quotient Lenses

### 2018-03-16 papers/QuotientLenses/main.tex
Response about colors

\newif\ifdraft\drafttrue  % set true to show comments
\newif\iflastminute\lastminutetrue  % for things to check at the end

@@

% %%%% Macros Colors
\definecolor{dkblue}{rgb}{0,0.1,0.5}
\definecolor{dkgreen}{rgb}{0,0.6,0}
\definecolor{dkred}{rgb}{0.6,0,0}
\definecolor{dkpurple}{rgb}{0.7,0,0.4}
\definecolor{olive}{rgb}{0.4, 0.4, 0.0}
\definecolor{teal}{rgb}{0.0,0.5,0.5}
\definecolor{orange}{rgb}{0.9,0.6,0.2}
\definecolor{lightyellow}{RGB}{255, 255, 179}
\definecolor{lightgreen}{RGB}{170, 255, 220}
\definecolor{teal}{RGB}{141,211,199}
\definecolor{darkbrown}{RGB}{121,37,0}

@@

This paper applies this idea to the problem of
synthesizing {\em quotient lenses}~\cite{quotientlenses}.
Quotient lenses are lenses in which the lens laws are
loosened so that they hold modulo an equivalence relation on the source and
target data respectively; in this paper we are concerned with {\em bijective
quotient lenses} which are lenses for which Equation \ref{bijectivelenslaws}
holds modulo equivalence relations $\equiv_S$ and $\equiv_T$ defined on the source
and target data respectively:\bcp{Not too convinced about the colors.  The
  fact that lens expressions mix black and red is distracting -- my eyes
  can't easily tell which bits of a larger expression are lenses.  I would
  suggest either:
  \begin{itemize}
  \item get and put italicized, plus
  \item either and lens expressions (completely) in some subtle color (maybe
  brown), or simply black.
  \end{itemize}
}\saz{I agree that the coloring can be a bit distracting, but I do think that
  highlighting the keywords in the code helps for parsing--maybe using dark blue
  is better? Why do ``get'' and ``put'' deserve special treatment as keywords,
  compared to, say, ``synth'' or ``del''?}
\begin{equation}\label{quotientlenslaws}
\ell.\get \; (\ell.\lput \; t) \equiv_T t \text{, and } \ell.\lput \; (\ell.\get
\; s) \equiv_S s
\end{equation}
Quotient lenses are useful in situations where a programmer wishes for the
transformation defined by a lens to have the same behavior on data that differ
only in inessential details. For instance, a programmer may wish to ``quotient
away'' the number of white space characters between data items, the
capitalization of various strings, the sequence of fields in a record, or the
order of items in a list.

@@

Indeed this idea is what motivates our definition of {\em QRE lenses}.
Intuitively, our QRE lenses are bijective lenses with ``canonizers at the
edges''.
Figure~\ref{fig:attheedges} depicts the architecture of QRE lenses.
\begin{figure}[t]
\centering
\includegraphics[width=0.7\textwidth]{QREs}
\caption{QRE lens from source $P$ to target $Q$.  $P$ and $Q$ are QREs, each
  consisting of a ``whole'' set ($W(P)$ and $W(Q)$) and a ``kernel'' part
  ($K(P)$ and $K(Q)$) Between the two kernels is a plain bijective lens.
  The \kw{get} function of the whole lens takes an argument from $W(P)$,
  applies the canonizer for $P$ to obtain a canonical representative in
  $K(P)$, then applies the \kw{get} of the plain lens, yielding an element
  of $K(Q)$, which is a subset of $W(Q)$.  The \kw{put} function does the
  reverse, mapping from $W(Q)$ to $K(P)$ (hence $W(P)$.  This lens is in
  normal form, with canonizers (specified by QREs) at the outer edges and an
  ordinary bijective lens in the center. \BCP{Both the figure and the
    caption are new---what do you think??}}\saz{I think that this figure is a
  big improvement.  One tiny possible tweak: color the \kw{get} and \kw{put}
  keywords in the same color as in the text.}\bcp{Can do, if we keep the
  color in the text :-).  How are we going to decide about that?}\saz{Vote?}
\label{fig:attheedges}
\end{figure}
Every QRE lens $q$ has a type $P \Leftrightarrow Q$ where $P$ and $Q$
are QREs. In the {\em get} direction, a QRE lens $q: P \Leftrightarrow Q$ uses
the source QRE $P$ to compute a canonical representative for the data modulo
the equivalence relation defined by $P$ and then applies the \kw{get} function
of a bijective string lens $\ell$ to this representative. In the {\em put}
direction, $q$ operates similarly, but using the QRE $Q$ and the $\lput$
function of $\ell$.

>>>>

\newif\ifdraft\drafttrue  % set true to show comments
\newif\iflastminute\lastminutetrue  % for things to check at the end
\newif\iflater\laterfalse  % for things to think about after submission

@@

% %%%% Macros Colors
\definecolor{dkblue}{rgb}{0,0.1,0.7}
\definecolor{dkgreen}{rgb}{0,0.6,0}
\definecolor{dkred}{rgb}{0.6,0,0}
\definecolor{dkpurple}{rgb}{0.7,0,0.4}
\definecolor{olive}{rgb}{0.4, 0.4, 0.0}
\definecolor{teal}{rgb}{0.0,0.5,0.5}
\definecolor{orange}{rgb}{0.9,0.6,0.2}
\definecolor{lightyellow}{RGB}{255, 255, 179}
\definecolor{lightgreen}{RGB}{170, 255, 220}
\definecolor{teal}{RGB}{141,211,199}
\definecolor{darkbrown}{RGB}{121,37,0}

@@

This paper applies this idea to the problem of
synthesizing {\em quotient lenses}~\cite{quotientlenses}.
Quotient lenses are lenses in which the lens laws are
loosened so that they hold modulo an equivalence relation on the source and
target data respectively; in this paper we are concerned with {\em bijective
quotient lenses} which are lenses for which Equation \ref{bijectivelenslaws}
holds modulo equivalence relations $\equiv_S$ and $\equiv_T$ defined on the source
and target data respectively:\bcp{Not too convinced about the colors.  The
  fact that lens expressions mix black and red is distracting -- my eyes
  can't easily tell which bits of a larger expression are lenses.  I would
  suggest either:
  \begin{itemize}
  \item get and put italicized, plus
  \item either and lens expressions (completely) in some subtle color (maybe
  brown), or simply black.
  \end{itemize}
}\saz{I agree that the coloring can be a bit distracting, but I do think that
  highlighting the keywords in the code helps for parsing--maybe using dark blue
  is better? Why do ``get'' and ``put'' deserve special treatment as keywords,
  compared to, say, ``synth'' or ``del''?}
%
\bcp{This gets at the basic difficulty with code coloring: we are dealing,
  in different parts of the paper, with (1) Boomerang code and (2)
  mathematical expressions involving lenses. We might want to make different
  highlighting choices in the two cases (e.g., we might want to highlight
  just individual keywords in large pieces of code, rather than the whole
  thing, while lens expressions within larger math formulas should perhaps
  be colored all the same, to better distinguish them from their context.
  My suggestion for get and put was based on the fact that they are not
  really part of boomerang programs: they are only mathematical entities (I
  think you {\em can} extract them as separate functions in boomerang, but
  usually there is no reason to, and we don't do it here).  However, I think
  it doesn't matter that much and we should use our remaining time now for
  polishing the text.  I propose we try dark blue for today and, if it's not
  distracting, just leave it alone.}
%
\begin{equation}\label{quotientlenslaws}
\ell.\get \; (\ell.\lput \; t) \equiv_T t \text{, and } \ell.\lput \; (\ell.\get
\; s) \equiv_S s
\end{equation}
Quotient lenses are useful in situations where a programmer wishes for the
transformation defined by a lens to have the same behavior on data that differ
only in inessential details. For instance, a programmer may wish to ``quotient
away'' the number of white space characters between data items, the
capitalization of various strings, the sequence of fields in a record, or the
order of items in a list.

@@

Indeed this idea is what motivates our definition of {\em QRE lenses}.
Intuitively, our QRE lenses are bijective lenses with ``canonizers at the
edges''.
Figure~\ref{fig:attheedges} depicts the architecture of QRE lenses.
\begin{figure}[t]
\centering
\includegraphics[width=0.7\textwidth]{QREs}
\caption{QRE lens from source $P$ to target $Q$.  $P$ and $Q$ are QREs, each
  consisting of a ``whole'' set ($W(P)$ and $W(Q)$) and a ``kernel'' part
  ($K(P)$ and $K(Q)$) Between the two kernels is a plain bijective lens.
  The \kw{get} function of the whole lens takes an argument from $W(P)$,
  applies the canonizer for $P$ to obtain a canonical representative in
  $K(P)$, then applies the \kw{get} of the plain lens, yielding an element
  of $K(Q)$, which is a subset of $W(Q)$.  The \kw{put} function does the
  reverse, mapping from $W(Q)$ to $K(P)$ (hence $W(P)$.  This lens is in
  normal form, with canonizers (specified by QREs) at the outer edges and an
  ordinary bijective lens in the center. }
\iflater
\saz{I think that this figure is a
  big improvement.  One tiny possible tweak: color the \kw{get} and \kw{put}
  keywords in the same color as in the text.}\bcp{Can do, if we keep the
  color in the text :-).  How are we going to decide about that?}\saz{Vote?}
\fi
\label{fig:attheedges}
\end{figure}
Every QRE lens $q$ has a type $P \Leftrightarrow Q$ where $P$ and $Q$
are QREs. In the {\em get} direction, a QRE lens $q: P \Leftrightarrow Q$ uses
the source QRE $P$ to compute a canonical representative for the data modulo
the equivalence relation defined by $P$ and then applies the \kw{get} function
of a bijective string lens $\ell$ to this representative. In the {\em put}
direction, $q$ operates similarly, but using the QRE $Q$ and the $\lput$
function of $\ell$.
